Examining background gene lists in clusterProfiler - RNA-seq example

### 2023-08-30 analysis/example_proteomics.Rmd
prot

For this report I will be using proteomics data from Mackmull et al (2015) where they looked at the
effect of HDACi on protein expression in nuclear and cytosolic compartments.
Here I am using the data from the TSA treated nuclear fractions after 48 hrs exposure.
The data were obtained from Supplementary Table 2 and saved to the /data folder.

@@

def <- read.table("../data/tsa_48hr_nuclear_frac.tsv",sep="\t",header=TRUE)

@@

| Reporting criteria | Method/resource used |
| --- | --- |
| Origin of gene sets | KEGG (2023-06-16) |
| Tool used | ClusterProfiler (check version at foot of report) |
| Statistical test used | hypergeometric test |
| P-value correction for multiple comparisons | FDR method |
| Background list | Genes with >=10 reads per sample on average across all samples |
| Gene Selection Criteria | DESeq2 FDR<0.05 |
| ORA directionality | Separate tests for up- and down-regulation |
| Data availability | Mackmull et al (2015) |
| Other parameters | Min gene set size of 10 |

@@

defup <- unique(subset(def,adj.P.Val<=0.05 & logFC>0)$GN)

defdn <- unique(subset(def,adj.P.Val<=0.05 & logFC<0)$GN)

bg <- unique(def$GN)

@@

| Reporting criteria | Method/resource used |
| --- | --- |
| Origin of gene sets | KEGG (2023-06-16) |
| Tool used | ClusterProfiler (check version at foot of report) |
| Statistical test used | hypergeometric test |
| P-value correction for multiple comparisons | FDR method |
| Background list | Genes with >=10 reads per sample on average across all samples |
| Gene Selection Criteria | DESeq2 FDR<0.05 |
| ORA directionality | Separate tests for up- and down-regulation |
| Data availability | Mackmull et al (2015) |
| Other parameters | Min gene set size of 10 + BUG FIX |

@@

| Reporting criteria | Method/resource used |
| --- | --- |
| Origin of gene sets | Reactome (2023-03-06) |
| Tool used | ClusterProfiler (check version at foot of report) |
| Statistical test used | hypergeometric test |
| P-value correction for multiple comparisons | FDR method |
| Background list | Genes with >=10 reads per sample on average across all samples |
| Gene Selection Criteria | DESeq2 FDR<0.05 |
| ORA directionality | Separate tests for up- and down-regulation |
| Data availability | Mackmull et al (2015) |
| Other parameters | Min gene set size of 10 |

@@

| Reporting criteria | Method/resource used |
| --- | --- |
| Origin of gene sets | Reactome (2023-03-06) |
| Tool used | ClusterProfiler (check version at foot of report) |
| Statistical test used | hypergeometric test |
| P-value correction for multiple comparisons | FDR method |
| Background list | Genes with >=10 reads per sample on average across all samples |
| Gene Selection Criteria | DESeq2 FDR<0.05 |
| ORA directionality | Separate tests for up- and down-regulation |
| Data availability | Mackmull et al (2015) |
| Other parameters | Min gene set size of 10 + BUG FIX |

>>>>

For this report I will be using proteomics data from Stukalov et al (https://pubmed.ncbi.nlm.nih.gov/33845483/),
who profiled the effect of SARS-COV2 infection on A549 cells.

Here we used the 24hr profile, as there were relatively more changes found.

Probable contaminants were removed, as well as viral proteins.
This left 5825 proteins, including 165 downregulated and 44 upregulated.

The data were obtained from Supplementary Table 5 and saved to the /data folder.

@@

def <- read.table("../data/sarscov2_prot.tsv",sep="\t",header=TRUE)

@@

| Reporting criteria | Method/resource used |
| --- | --- |
| Origin of gene sets | KEGG (2023-06-16) |
| Tool used | ClusterProfiler (check version at foot of report) |
| Statistical test used | hypergeometric test |
| P-value correction for multiple comparisons | FDR method |
| Background list | Genes with >=10 reads per sample on average across all samples |
| Gene Selection Criteria | DESeq2 FDR<0.05 |
| ORA directionality | Separate tests for up- and down-regulation |
| Data availability | Stukalov et al 2021: PMID 33845483 |
| Other parameters | Min gene set size of 10 |

@@

defup <- unique(subset(def,change.SARS_CoV2.24h_vs_mock.24h=="+")$gene_name)

defdn <- unique(subset(def,change.SARS_CoV2.24h_vs_mock.24h=="-")$gene_name)

bg <- unique(def$gene_name)

@@

| Reporting criteria | Method/resource used |
| --- | --- |
| Origin of gene sets | KEGG (2023-06-16) |
| Tool used | ClusterProfiler (check version at foot of report) |
| Statistical test used | hypergeometric test |
| P-value correction for multiple comparisons | FDR method |
| Background list | Genes with >=10 reads per sample on average across all samples |
| Gene Selection Criteria | DESeq2 FDR<0.05 |
| ORA directionality | Separate tests for up- and down-regulation |
| Data availability | Data availability | Stukalov et al 2021: PMID 33845483 |
| Other parameters | Min gene set size of 10 + BUG FIX |

@@

| Reporting criteria | Method/resource used |
| --- | --- |
| Origin of gene sets | Reactome (2023-03-06) |
| Tool used | ClusterProfiler (check version at foot of report) |
| Statistical test used | hypergeometric test |
| P-value correction for multiple comparisons | FDR method |
| Background list | Genes with >=10 reads per sample on average across all samples |
| Gene Selection Criteria | DESeq2 FDR<0.05 |
| ORA directionality | Separate tests for up- and down-regulation |
| Data availability | Data availability | Stukalov et al 2021: PMID 33845483 |
| Other parameters | Min gene set size of 10 |

@@

| Reporting criteria | Method/resource used |
| --- | --- |
| Origin of gene sets | Reactome (2023-03-06) |
| Tool used | ClusterProfiler (check version at foot of report) |
| Statistical test used | hypergeometric test |
| P-value correction for multiple comparisons | FDR method |
| Background list | Genes with >=10 reads per sample on average across all samples |
| Gene Selection Criteria | DESeq2 FDR<0.05 |
| ORA directionality | Separate tests for up- and down-regulation |
| Data availability | Data availability | Stukalov et al 2021: PMID 33845483 |
| Other parameters | Min gene set size of 10 + BUG FIX |

### 2023-08-30 analysis/example_rnaseq.Rmd
prot

png("rna_kegg_euler")
plot(euler(v3),quantities = list(cex = 2), labels = list(cex = 2),main="up- and down-regulated KEGG pathways")
dev.off()

@@

png("rna_reactome_euler")
plot(euler(v3),quantities = list(cex = 2), labels = list(cex = 2),main="up- and down-regulated Reactomes")
dev.off()

>>>>

png("rna_kegg_euler.png")
plot(euler(v3),quantities = list(cex = 2), labels = list(cex = 2),main="up- and down-regulated KEGG pathways")
dev.off()

@@

png("rna_reactome_euler.png")
plot(euler(v3),quantities = list(cex = 2), labels = list(cex = 2),main="up- and down-regulated Reactomes")
dev.off()
